Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers

### 2021-07-08 docs/manuscript.Rmd
Update abstract, intro, methods

---
title: "Linkage of trial registrations and results publications: A cross-sectional, automated evaluation of German university medical centers"
author: "Maia Salholz-Hillel, Daniel Strech, Benjamin Gregory Carlisle"
output: 
  bookdown::word_document2:
    reference_docx: quest-word-template.docx
    latex_engine: xelatex
bibliography: references.bib
editor_options:
    markdown:
        canonical: true
    chunk_output_type: console
---

@@

**Objective** To evaluate, using a novel automated approach, the linkage of registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). The International Committee of Medical Journal Editors (ICMJE) and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linkage to promote discoverability of both the clinical trial publication and registration and thus increase the value of clinical research.

@@

**Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered clinical trials conducted across German UMCs and completed between 2009 and 2017. We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and TRNs in publication metadata, abstract, and full-text. <!--# We verified our automated approach on a manually coded sample. -->

@@

**Secondary analyses** We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and publication linkage in the registration. We also explored the relationship between cross-linkage<!--# practices --> and registry. We also explored the relationship between cross-linkage<!--# practices --> and prospective registration <!--# not sure about this. i think if we do this then glm for the various practices -->. We looked at change in registration-publication linkage over time. We also looked at cross-registrations in the registries as well as additional TRNs reported in trial publications.<!--#  We also looked at reporting PMC-version full-text. We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status. We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting. All analyses are exploratory. -->

**Results** Of `r n_trials_ft_pdf` trials with PubMed-indexed clinical trial results publications available as full-text, we found `r report_n_pct_trials(n_trials_link_full)` trials with comprehensive registration-publication linkage, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links. We found that only `r report_n_pct_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, whereas `r report_n_pct_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials did so, mostly via automatic publication indexing. TRNs were most commonly reported in the full-text, appearing in `r report_n_pct_trials(n_trials_link_trn_ft)`. Linkage rates appear overall lower for DRKS trials than for ClinicalTrials.gov. The most common link is TRN reporting in the full-text which reached `r stringr::str_extract(report_link_year("TRN (Full-text)", 2017), "\\d{2}%")` in 2017 in both registries. <!--# add more secondary? -->

**Conclusion** Despite encouragement from CONSORT and ICMJE, German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications. Publication linkage was noticeably greater in ClinicalTrials.gov which uses a more structured data entry process as well as automatically indexes publications from PubMed. German and international university medical centers (UMCs) should incentivize, promote, and educate their researchers on this responsible research practice. Automated strategies, such as those used in this study, could support bibliometric databases and registries to enrich their registration-publication linkage. For example, to improve TRN inclusion in its metadata, PubMed could automatically extract TRNs from publication abstracts or full-text. DRKS could adopt strategies like those used by ClinicalTrials.gov to automatically index publications reporting the TRN in PubMed. As compared with other responsible research practices (such as trial registration itself), reporting TRN and publications are copy-and-paste tasks that may be integrated into biomedical researchers' existing workflows.

@@

Clinical trial registration and results reporting are important tools for clinical research transparency [@worldmedicalassociation2013; @worldhealthorganization2017].<!--# I removed background on registration and results reporting for succinctness and because there's a robust literature, which I can add additional citations to. Let me know if you think it's worth adding back a few sentence of background. --> Linkage of the trial registrations and results publications promotes discoverability of both components <!--# adds further transparency beyond registration and results reporting independently -->and increases the value of clinical research. Such threaded evidence <!--# Linked registrations-publications-->empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd]. Robust linkage includes both the incorporation of a structured reference to publications in the registration, and reporting of trial registration numbers (TRN) in the full-text, abstract, and metadata of publications.

The Consolidated Standards of Reporting Trials (CONSORT) as well as the ICMJE's *Recommendations for the Conduct, Reporting, Editing, and Publication of Scholarly Work in Medical Journals* call for reporting "trial registration number and name of the trial register" in both the full-text and abstract of trial results publication [@internationalcommitteeofmedicaljournaleditorsicmje2019; @hopewell2008pm]. Reporting TRN solely in the full-text does not suffice since readers may not have access to the full-text or may base their judgment of the trial or topic on the abstract alone [@hopewell2008tl]. Abstracts may also be published independently of a full-text publication, such as for conference abstracts, highlighting the need for TRN reporting in abstracts as well [@hopewell2008pm].

Furthermore, the inclusion of TRNs in publication and bibliographic metadata enhances discoverability for both human readers and computerized analysis. For example, PubMed MEDLINE metadata includes TRNs under the "secondary source id" tag [@nationallibraryofmedicine2018], which can then be searched on the web interface (e.g., "NCT00002568[secondary identifier]" or "NCT00002568[si]") or accessed programmatically via the underlying XML [@nationallibraryofmedicine2017c]. Currently, these tags can be generated from publisher-provided data or assigned manually by NLM Indexing Data Reviewers who copy-and-paste secondary source ids (including trial ids) found in the abstract and full-text (personal communication via NLM Helpdesk Ticket CAS-552810-T3H7V5). <!--# @daniel: this had been in the intro but you said to shorten so i moved to discussion but then you had several quetion about how SI is made, so moved it back to intro... -->

Machine-readable TRN metadata moreover facilitates automatic incorporation of the publication in the registry for "easy linking of trial related publications with clinical trial registry site records" [@nationallibraryofmedicine2017c]. For example, ClinicalTrials.gov uses PubMed secondary identifiers to automatically index publications in the registry entry under "Additional publications automatically indexed to this study by ClinicalTrials.gov Identifier (NCT Number)" [@nationallibraryofmedicine2017c]. This therefore increases the discoverability of the publication from the registration, with no additional effort from the trial investigator for manually adding the publication to the registry.

Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings. <!--# add policy background on why pub in reg --> The insertion and structuring of references depends on the registry; whereas ClinicalTrials.gov includes both automatically indexed publications and manual entries in quasi-structured datafields, DRKS currently accepts only manual entries in minimally structured datafields. References, like those automatically indexed, may include one or more publication unique identifier such as a DOI or PMID, or may be limited to a free-text citation with journal, article, and author names, but no unique identifier. While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval and linkage [@huser2013po].

Trial evidence appears to be insufficiently linked in both registries and publications. @huser2013po found that only 28% of 8907 completed trials contained a structured trial-article link with a publication unique identifier in the registry. In a systematic review of 16 studies examining cohorts of trial publications for links to registrations<!--# of results publications with associated registrations/investigating linked registrations-publications in publication cohorts found via both automated and manual approaches...couldn't find the overall TRN reporting rate from all studies looking at TRN, only the 16 using both automated and manual approaches, but this level of detail feels like too much to explain -->, @bashir2017sr found an overall rate of 49% of results publications <!--# (range: 8-97%) --> reported a TRN in either the metadata, abstract, or full-text. More recently, @carlisle2020m found 38,001 ClinicalTrials.gov ids across 488,365 abstracts from records indexed by PubMed as clinical trial publications, meaning a maximum of `r round(38001/488365 * 100, digits = 1)`% of publications report a ClinicalTrials.gov id, assuming one TRN per publication. @aldurra2020b found that only 28.8% (7473/10,500) records classified as RCTs based on a bespoke PubMed query reported trial ids in any of the metadata, abstract, or full-text. <!--# @murph sanity check of numbers from other papers -->

These previous projects, however, faced limitations due to the unclear publication cohort and unknown true registration rates. PubMed queries currently do not suffice to identify the true populations of clinical trials or even RCTs, both excluding true positives (i.e., sensitivity) and including false positives (i.e., specificity) [@glanville2020jmla]. Therefore not all PubMed-indexed clinical trials or RCTs should be registered <!--# as they may not reflect true clinical trials or RCTs -->, and not all that should be are. Furthermore, not all reported TRNs correspond to bona fide registrations of trial whose results are reported in the publication, i.e., TRNs may be reported for related trials in the background section. As such, conclusions drawn by these previous studies relying on PubMed-query based samples, about registration reporting rates must be qualified as the rate of reported trial registration numbers in PubMed-indexed (randomized) clinical trials with unknown registration rates. <!--# remove previous sentence? --> In addition, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link. These studies of publication linkage in registrations have furthermore focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of other studies investigating publication linkage in DRKS.

@@

We evaluated the linkage of registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017. <!--# @daniel comment: add reference to intovalue here? it's in methods -->We also looked at change in these linkage practices over time. By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication linkage, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata. We use an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs. We did not pre-specify the analyses.

@@

## Data sources and collection

@@

We <!--# programmatically -->downloaded data from both registries on 15 May 2021. We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data. We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti]. As DRKS does not provide an API, we built a webscraper to capture the necessary fields [@federalinstitutefurdrugsandmedicaldevicesbfarm]. ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information. For example, ClinicalTrials.gov has a separate summary results reporting section, whereas summary results are included with other references in DRKS. We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse"). <!--# remove if not including summary results -->

We then queried the PubMed Entrez Programming Utilities (E-utilities) API via an R client for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter]. From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata, as well as MeSH terms, keywords, author details, and summary metadata. <!--# remove unused pubmed components? -->

We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies? --> [@ourresearch; @jahn]. We then used the GROBID machine learning library for technical and scientific publications to parse the PDFs into machine-readable XMLs [@2008]. In order to isolate the main body from the publication abstract for our analysis, we extracted the `<body>` sections of the papers.

Figure \@ref(fig:flow-methods) show the methods flow using automated techniques with the exception of some full-text PDFs.

@@

As reference data is available in difference formats depending on the registry, we developed parallel approaches to extract publication identifiers for each. For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT. We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field. For DRKS, we scraped the reference type, citation, and URL link, when available. We used regexes to extract any DOIs and PMIDs from either the citation or link. In the rare cases when conflicting DOIs or PMIDs were found in the citation vs. the link, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation. In addition, we determined whether a reference was manually or automatically indexed in the registry. DRKS allows only manually added references; for ClinicalTrials.gov, references of type "derived" were marked as automatically indexed. We then used DOIs and PMIDs to match the referenced publications to those in our trial dataset. We did not attempt to match publications without identifiers (i.e., publication title only).

@@

We developed <!--# and evaluated [or more methods details] --> regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(trn-regular-expressions) Table \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text. To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata for true positives and true negatives.<!--# In order to isolate the full-text main body from the publication abstract for our analysis, we included only `<body>` sections of the PMC and PDF samples for the full-text source. --> For each source, we first extracted all unique TRN patterns within a source. Our regex allowed for minor formatting errors in the TRN patterns (such as additional punctuation). We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source introduced through the cleaning process. We also coalesced across sources to deduce a list of unique TRNs by publication along with the source(s) in which each was reported.

@@

<!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration. currently placeholder "TRUE" such that all TRNs resolve in the registries -->

<!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns? then would have to change my backend work. or can keep this section but maybe not report all the details in text but just appendix? -->

@@

Data collection, preparation, and analysis was performed in R version 4.0.2 (R Foundation for Statistical Computing) and RStudio version 1.4 (RStudio Team). PDF conversion to TEI XML was performed using GROBID Server 0.6.1 [@2008] and GROBID Python Client [@lopez2021]. All code and the final analysis data for this study are available at [insert link]. Raw data (with the exception of the full-text of publications) is available at [insert link].

@@

The IntoValue trials dataset included the full sample of trials led by a German UMC that were registered on ClinicalTrials.gov or the DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria <!--# excluding non-German UMC-led trials, deduplicating trials, excluding trials without results, enhancing the publication metadata, getting full-text, and excluding trials with results without a PMID and full-text -->, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications. <!--# QUESTION: what summary info about trials and publications? I also include a long version of demographic data in case we want to include any of these, ideally with collapsed categories, in Table @ref(tab:tbl-trials-full) -->

@@

## Registration-publication linkage

While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linkage, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication. Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the breakdown of registration-publication linkage across the various components for ClinicalTrial.gov and DRKS trials.

```{r plot-upset-links-reg-pub-registries, fig.cap = "Registristion-publication linkages in ClinicalTrials.gov and DRKS", out.width="80%"}
knitr::include_graphics(here("docs", "figures", "plot-upset-links-reg-pub-registry.png"))
```

@@

We found that `r report_pct_num_denom_trials(n_trials_link_reg)` registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results. Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed.

@@

### Linkage by registry

We found that trials with a ClinicalTrials.gov TRN were significantly more likely than trials with a DRKS TRN to include a publication reference in the registry (`r apa::apa(chisq_link_by_reg_reg, print_n = TRUE)`) and to include the TRN in the PubMed metadata (`r apa::apa(chisq_link_by_reg_si, print_n = TRUE)`) and abstract (`r apa::apa(chisq_link_by_reg_abs, print_n = TRUE)`). We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`). Table \@ref(tab:tbl-link-by-reg) shows summary registration-publication linkage by registry.

@@

```{r tbl-link-by-reg}
tbl_link_by_reg %>% 
  gtsummary::as_flex_table() %>% 
  flextable::set_caption(caption = "Registration-publication linkage by registry")
```

@@

Linkage rates appear overall lower for DRKS trials than for ClinicalTrials.gov. Overall, we observed positive trends for TRN reporting in full-text and abstracts. <!--# add test? --> Figure \@ref(fig:plot-links-year-reg) shows the registration-publication linkages over time for ClinicalTrials.gov and DRKS.

>>>>

---
title: "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers"
author: "Maia Salholz-Hillel, Daniel Strech, Benjamin Gregory Carlisle"
output: 
  bookdown::word_document2:
    reference_docx: quest-word-template.docx
    latex_engine: xelatex
    fig_caption: true
    number_sections: false
bibliography: references.bib
editor_options:
    markdown:
        canonical: true
    chunk_output_type: console
---

@@

**Objective** To evaluate, using a novel automated approach, the links between registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to promote discoverability of both clinical trial publications and registrations and thus increase the value of clinical research.

@@

**Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017. We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text. <!--# We verified our automated approach on a manually coded sample. -->

@@

**Secondary analyses** We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and a linked publication in the registration. We also explored the relationship between registration-publication links and (1) registry, and (2) prospective registration <!--# not sure about this. i think if we do this then glm for the various practices -->. We looked at change in registration-publication linking practices over time. We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.<!--# We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status. We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting. All analyses are exploratory. -->

**Results** Only `r report_pct_num_denom_trials(n_trials_link_full)` of trials comprehensively link their registrations and publications, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, `r report_n_pct_trials(n_trials_link_none)` trials had no links. The most common link was TRN reported in the full-text, which occurred for `r report_n_pct_trials(n_trials_link_trn_ft)` trials <!--# and reached `r stringr::str_extract(report_link_year("TRN (Full-text)", 2017), "\\d{2}%")` in 2017 in both registries. --><!--# add more secondary? -->

DRKS trials were overall more poorly linked than ClinicalTrials.gov trials. For example, links to results publications appeared in `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov registrations, but only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS registrations. Most [`r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)`] of the links in ClinicalTrials.gov were generated via automatic publication indexing from PubMed metadata, which DRKS does not do.

**Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications. ClinicalTrials.gov trials were more extensively linked and rely on a more structured data entry process and automatic publication indexing from PubMed; DRKS could adopt similar strategies to improve linking.

German and international university medical centers (UMCs) should incentivize, promote, and educate their researchers on this responsible research practice. Automated strategies, such as those used in this study, could support bibliometric databases and registries to enrich their registration-publication links. For example, to improve TRN inclusion in its metadata, PubMed could automatically extract TRNs from publication abstracts or full-text. As compared with other responsible research practices (such as trial registration itself), reporting TRN and publications are copy-and-paste tasks that may be integrated into biomedical researchers' existing workflows.

@@

Clinical trial results risk being left out of evidence synthesis and health policy decision making if publications cannot be found. Incomplete evidence threatens healthcare and the lives and welfare of human volunteers. Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp]. Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing the clinical research value[@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->. Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].

Trial results should be bidirectionally linked, meaning the inclusion of both a structured reference to publications in the registration, and trial registration numbers (TRN) in the full-text, abstract, and metadata of publications. The Consolidated Standards of Reporting Trials (CONSORT) as well as the ICMJE's *Recommendations for the Conduct, Reporting, Editing, and Publication of Scholarly Work in Medical Journals* call for reporting "trial registration number and name of the trial register" in both the full-text and abstract of trial results publication [@internationalcommitteeofmedicaljournaleditorsicmje2019; @hopewell2008pm]. Reporting TRN solely in the full-text does not suffice since readers may not have access to the full-text or may base their judgment of the trial or topic on the abstract alone [@hopewell2008tl]. Abstracts may also be published independently of a full-text publication, such as for conference abstracts, highlighting the need for TRN reporting in abstracts as well [@hopewell2008pm].

The inclusion of TRNs in publication and bibliographic metadata especially enhances machine readability and discoverability. Trial results can then be found more efficiently using TRNs to search publication databases [@huser2013po]. Publications can then also be automatically linked within trial registrations, as is currently done with ClinicalTrials.gov using TRN metadata from PubMed [@nationallibraryofmedicine2017c]. Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings. <!--# add policy background on why pub in reg --> Results references may be entered manually in many registries, with varying degrees of structure. ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas DRKS offers an unstructured free-text field. <!--# add some background on ctgov vs. drks. and how they include trns and publications. and ctgov and pubmed by same nlm so may be better at seing those ids -->While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval [@huser2013po].

Trial evidence appears to be insufficiently linked in both registries and publications. Previous studies on trial registrations found links to results publications in as few as 13% to as many as 42% of trials [@bashir2017sr; @huser2013po]. TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b]. For example, some studies included trial publications limited to specific journals or fields whereas others relied on PubMed "clinical trial" publication type queries.

<!--# REMOVE or @murph sanity check of numbers from other papers -->

@huser2013po found that only 28% of 8907 completed trials contained a structured trial-article link with a publication unique identifier in the registry. In a systematic review of 16 studies examining cohorts of trial publications for links to registrations<!--# of results publications with associated registrations/investigating linked registrations-publications in publication cohorts found via both automated and manual approaches...couldn't find the overall TRN reporting rate from all studies looking at TRN, only the 16 using both automated and manual approaches, but this level of detail feels like too much to explain -->, @bashir2017sr found that a median of 49% of results publications (range: 8-97%) reported a TRN in either the metadata, abstract, or full-text. More recently, @carlisle2020m found 38,001 ClinicalTrials.gov ids across 488,365 abstracts from records indexed by PubMed as clinical trial publications, meaning a maximum of `r round(38001/488365 * 100, digits = 1)`% of publications report a ClinicalTrials.gov id, assuming one TRN per publication. @aldurra2020b found that only 28.8% (7473/10,500) of publications classified as RCTs based on a bespoke PubMed query reported trial ids in any of the metadata, abstract, or full-text.

In order to draw robust conclusions about registration-publication linking, the analysis population must be limited to registered clinical trials with published results. Previous projects faced limitations due to the unclear publication cohort and unknown true registration rates. PubMed queries currently do not suffice to identify the true populations of clinical trials or even RCTs, both excluding true positives (i.e., sensitivity) and including false positives (i.e., specificity) [@glanville2020jmla]. Therefore not all PubMed-indexed clinical trials or RCTs should be registered <!--# as they may not reflect true clinical trials or RCTs -->, and not all that should be are. Furthermore, not all reported TRNs correspond to bona fide registrations of the trial whose results are reported in the publication, i.e., TRNs may be reported for related trials in the background section. As such, conclusions drawn by these previous studies relying on PubMed-query based samples, about registration reporting rates must be qualified as the rate of reported trial registration numbers in PubMed-indexed (randomized) clinical trials with unknown registration rates. <!--# remove previous sentence? --> Similarly, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link. Furthermore, these studies of publication links in registrations have focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of other studies investigating publication links in DRKS.

@@

We evaluated linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017. We also looked at change in these practices over time. By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata. We use an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs. We did not pre-specify the analyses.

@@

## Automated pipeline for data sources and collection

@@

We <!--# programmatically -->downloaded data from both registries on 15 May 2021. <!--# We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data. -->We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti]. As DRKS does not provide an API, we built a webscraper to capture the necessary fields [@federalinstitutefurdrugsandmedicaldevicesbfarm]. <!--# ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information. For example, ClinicalTrials.gov has a separate summary results reporting section, whereas summary results are included with other references in DRKS. We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse").-->

We queried the PubMed Entrez Programming Utilities (E-utilities) API for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter]. From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata, as well as MeSH terms, keywords, author details, and summary metadata. <!--# remove unused pubmed components? -->

We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies? -->[@ourresearch; @jahn]. We then used the GROBID machine learning library for technical and scientific publications (v0.6.1) via Python to parse the PDFs into machine-readable XMLs [@2008; @lopez2021]. In order to isolate the main body from the publication abstract for our analysis, we extracted the `<body>` sections of the papers.

Figure \@ref(fig:flow-methods) show the methods flow using automated techniques<!--# with the exception of some full-text PDFs -->.

@@

As reference data is available in difference formats depending on the registry, we developed parallel approaches to extract publication identifiers for each. For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT. We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field. For DRKS, we scraped the reference type, citation, and URL, when available. We used regexes to extract any DOIs and PMIDs from the citation and URL. In the rare cases when conflicting DOIs or PMIDs were found in the citation vs. the URL, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation. In addition, we determined whether a reference was manually or automatically indexed in the registry. DRKS allows only manually added references; for ClinicalTrials.gov, references of type "derived" were marked as automatically indexed. We then used DOIs and PMIDs to match the referenced publications to those in our trial dataset. We did not attempt to match publications without identifiers (i.e., publication title only).

@@

We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(trn-regular-expressions) Table \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text. To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata for true positives and true negatives. For each source, we first extracted all unique TRN patterns within a source. Our regex allowed for minor formatting errors in the TRN patterns (such as additional punctuation). We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source introduced through the cleaning process. We also coalesced across sources to deduce a list of unique TRNs by publication along with the source(s) in which each was reported.

@@

<!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration. currently placeholder "TRUE" such that all TRNs resolve in the registries --><!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns? then would have to change my backend work. or can keep this section but maybe not report all the details in text but just appendix? -->

@@

Data collection, preparation, and analysis was performed in R version 4.0.2 (R Foundation for Statistical Computing) and RStudio version 1.4 (RStudio Team). Additional packages and libraries are specified in the methods. All code and the final analysis data for this study are available at [insert link]. Raw data (with the exception of the full-text of publications) is available at [insert link].

@@

The IntoValue trials dataset included the full sample of trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications. <!--# QUESTION: what summary info about trials and publications? I also include a long version of demographic data in case we want to include any of these, ideally with collapsed categories, in Table @ref(tab:tbl-trials-full) -->

@@

## Registration-publication links

While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication. Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the breakdown of registration-publication links across the various components for ClinicalTrial.gov and DRKS trials.

```{r plot-upset-links-reg-pub-registries, fig.cap = "Registration-publication links in ClinicalTrials.gov and DRKS", out.width="80%"}
knitr::include_graphics(here("docs", "figures", "plot-upset-links-reg-pub-registry.png"))
```

@@

We found that `r report_pct_num_denom_trials(n_trials_link_reg)` of registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results. Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed.

@@

### Linking by registry

We found that trials with a ClinicalTrials.gov TRN were significantly more likely than trials with a DRKS TRN to include a publication reference in the registry (`r apa::apa(chisq_link_by_reg_reg, print_n = TRUE)`) and to include the TRN in the PubMed metadata (`r apa::apa(chisq_link_by_reg_si, print_n = TRUE)`) and abstract (`r apa::apa(chisq_link_by_reg_abs, print_n = TRUE)`). We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`). Table \@ref(tab:tbl-link-by-reg) shows summary registration-publication links by registry.

@@

```{r tbl-link-by-reg}
tbl_link_by_reg %>% 
  gtsummary::as_flex_table() %>% 
  flextable::set_caption(caption = "Registration-publication links by registry")
```

@@

Overall, DRKS trials appear to be less robustly linked than ClinicalTrials.gov trials. Overall, we observed positive trends for TRN reporting in full-text and abstracts. <!--# add test? --> Figure \@ref(fig:plot-links-year-reg) shows the registration-publication links over time for ClinicalTrials.gov and DRKS.
